### 2025-01-23 paper/paper.tex
Abstract has been writen and inserted image for quadtree.

\documentclass[a4paper, 12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{svg}

\renewcommand{\baselinestretch}{1.15}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\graphicspath{ {../graphs/} }

\begin{document}

\begin{titlepage}
    \centering
    \vspace*{1cm}

    {\Huge \textbf{Comparison of Spatial Partitioning Strategies}}\\[1.5cm]

    {\large Linus Kåreblom}\\[1cm]

    \vfill

    Technical program Academic year 2024/2025\\
    \textbf{School:} Österänggymnasiet\\
    \textbf{Class:} TE3A\\
    \textbf{Supervisor:} Tony Tieu\\

    \vspace{1cm}
    {\today}
\end{titlepage}

\begin{abstract}
    Spatial partitioning is a widely used optimization strategy to efficiently
    optimize objects in space. It can be found in many applications, from
    simulating galaxies to video games.
    Checking collision between objects is an expensive operation that quickly
    becomes a bottleneck as the number of objects grow.
    This paper compares three different strategies, alongside a naive solution.
    The aim is to evaluate the use cases for the different strategies.
    Each strategy was implemented to fit the same interface to make it as fair
    of a comparison as possible.
    The results show how spatial hashing out performs every other method when
    the objects are distributed evenly while quadtrees excel with uneven object
    distribution.
\end{abstract}

\newpage

\tableofcontents
\newpage

\section{Introduction}
When I was developing a small physics engine, I encountered performance issues
during the  collision detection phase. In this phase the program figures out
which objects are close enough to be touching and overlapping. This wouldn't be
an issue in a small simulation with few objects, but I was striving for
something grand running in real time which made every millisecond count. Due to
these constraints I had to find a way to speed things up.

On my optimization journey I encountered the concept of spatial partitioning -
an elegant solution to my problem. Diving deeper I discovered the three
techniques I'll be comparing in this paper, along side the naive solution. The
basic theory of the technique is simply to group objects close to each other to
avoid checking two objects on the opposite side of the world. This drastically
decreases the number of collision checks performed each frame.

As a consequence of this constraint I had to find a way to optimize this phase.
That is what sent me down the rabbit hole of spatial partitioning as an
optimization to do query objects close to each other in space. During my
research I found out about the three strategies I'm comparing in this paper. I'm
doing this paper within the programming course.

\section{Theory}
\subsection{Spatial Partitioning}
Spatial partitioning is an optimization technique used to make programs run
faster. It's often found within physics applications. The theory behind this
technique is to reduce the number of times one object is checked against
another. Suppose we have one hundred objects in our world. The naive way to
check each one of these against the others results in a ten thousand checks.
Since checks are expensive to make, we want to perform as few of them as
possible.

\subsection{Quadtree}
A quadtree is a recursive data structure, meaning it contains pointers to each
objects of the same type as itself. It's in the family of tree data structures.
The data structure is comprised of nodes. Each node has four pointers to other
nodes, could also be referred to as branches or leaves. This strategy works by
inserting points into a leaf until the leaf reaches its capacity, then it splits
into four equally large sub leaves. This subdivision pattern is shown in figure
\ref{fig:quadtree}.

\begin{figure}[ht]
    \includesvg[width=\textwidth]{images/invert-quadtree.svg}
    \caption{Visual representation of a quadtree.}
    \label{fig:quadtree}
\end{figure}

\subsection{The Stack and Heap}
A program uses two different types of memory, the stack and the heap. When the
program starts a fixed amount of bytes are allocated for the program to use,
this is what is called the stack. When the program creates a variable it is
allocated on the stack, which is done by incrementing the stack pointer by the
size of the variable data type. Since a stack allocation is only a single
instruction it is incredibly fast. The heap is much larger than the stack but
slower because you have to request space within it from the operating system.
Allocating is considered to be a slow operation since there's a lot of overhead
introduced.

\subsection{Big O notation}
Big O notation is a notation system to describe the time complexity of an
algorithm. If an algorithm has the time complexity of $O(n)$ then that means
its time complexity grows linearly with the amount of inputs $n$. This could be
deceiving due to a hidden constant within the expression. If $O(n)$ would be
written as a function to describe the actual execution time of the algorithm
then it would look like this $T(n) = C \cdot n$. A $O(n^{2})$ algorithm would
be considered worse than an $O(n)$ one. This assumption is, however, not always
correct. If $n$ is small enough then the $O(n^{2})$ algorithm could run faster
due to the constant being smaller than its $O(n)$ equivalent.

\subsection{Static Grid}
The static grid is built using a simple array. Indexing into the array is an
$O(1)$ operation. The caveat with this approach is you're forced to know how
big the area you want the to partition to be before compiling the program. Both
insertion and querying only requires a simple index into the grid array which
makes this approach super fast.

\subsection{Spatial Hashing Grid}
The spatial hashing grid is built using the same technique as a hash map. The
method I used to write my spatial hashing grid was to allocate an arbitrarily
sized array. When inserting into the structure I first hash the coordinate of
the object, then run a modulo operation on it giving me a number within the
array bounds, and then it is inserted into the cell.

\end{document}