### 2025-03-13 paper/paper.tex
paper: Final revision.

\documentclass[a4paper, 12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{svg}
\usepackage{float}
\usepackage{appendix}
\usepackage[hidelinks]{hyperref}

\hypersetup{
    colorlinks=false,
}

\renewcommand{\baselinestretch}{1.15}
\setlength{\parindent}{0pt}
\setlength{\parskip}{0.8em}

\graphicspath{ {../graphs/} }

\begin{document}

\pagenumbering{alph}
\begin{titlepage}
    \centering
    \vspace*{1cm}

    {\Huge \textbf{Spatial Partitioning Strategies}}\\[1.5cm]

    {\large Linus Kåreblom}\\[1cm]

    \vfill

    Technical program Academic year 2024/2025\\
    \textbf{School:} Österänggymnasiet\\
    \textbf{Class:} TE3A\\
    \textbf{Supervisor:} Tony Tieu\\

    \vspace{1cm}
    {\today}
\end{titlepage}
\pagenumbering{arabic}

\begin{abstract}
    Spatial partitioning is a widely used optimization strategy to efficiently
    optimize objects in space. It can be found in many applications, from
    simulating galaxies to video games. Checking collisions between objects is an
    expensive operation that quickly becomes a bottleneck as the number of
    objects grow. This paper compares three different strategies, alongside a
    naive solution. The aim is to evaluate the use cases for the different
    strategies. Each strategy was implemented to fit the same interface to make
    it as fair of a comparison as possible. The results show how quadtrees out
    performs every other method when the objects are evenly and uneven
    distributed as the number of objects grow.
\end{abstract}

\newpage

\tableofcontents
\newpage

\section{Introduction}
When I was developing a small physics engine, I encountered performance issues
during the  collision detection phase. In this phase the program figures out
which objects are close enough to be touching and overlapping. This would not be
an issue in a small simulation with few objects, but I was striving for
something grand running in real time which made every millisecond count. Due to
these constraints I had to find a way to speed things up.

On my optimization journey I encountered the concept of spatial partitioning -
an elegant solution to my problem. Diving deeper I discovered the three
techniques I will be comparing in this paper, along side the naive solution. The
basic theory of the technique is simply to group objects close to each other to
avoid checking two objects on the opposite side of the world. This drastically
decreases the number of collision checks performed each frame.

As a consequence of this constraint I had to find a way to optimize this phase.
That is what sent me down the rabbit hole of spatial partitioning strategies.
During my research I discovered the three strategies I am comparing in this
paper written under the subject of programming.

\section{Purpose \& Scope}
The purpose of this paper is to find the most optimal strategy for dividing up
space to improve the speed of spatial querying of nearby objects.

This study focuses mainly on CPU performance and runtime at the cost of higher
memory usage. Therefore time based evaluations of insertion, query time and
general collision testing time will be performed.

The strategies are going to be evaluated on time based metrics - insertion,
query and time spent on collision testing. Two object distributions are going
to be evaluated. A pseudo-random number generator will be used but given the
same seed of 1234 so all the pseudo-random numbers stay the same during every
measurement.

These strategies could be extended into the third dimension but this study
focuses solely on two dimensional space.

The questions I'm going to answer are:
\begin{itemize}
    \item At what point is it worth implementing a more advanced data structure?
    \item In what use case are the different algorithms most useful?
\end{itemize}

\section{Theory}
\subsection{Spatial Partitioning}
Spatial partitioning is an optimization technique used to make spatial queries
run faster. It is often found within physics applications. The theory behind
this technique is to reduce the number of object pairs being checked against
one another. Suppose we have one hundred objects in our world. The naive way to
make sure non of them are colliding is to check every object against every other
object in the world. This results in ten thousand collision checks. Since checks
are expensive, we want to perform as few of them as possible.

\subsection{Quadtree}
A quadtree is a recursive data structure, meaning it contains one or more
pointers, or references, to an object of the same type as one of its members. It
is in the family of tree data structures. The data structure is comprised of
nodes. Each node has four pointers to other nodes, which could also be referred
to as branches or leaves. This strategy works by inserting points into a leaf
until the leaf reaches max capacity, then it splits into four equally large sub
leaves. This subdivision pattern is shown in figure
\ref{fig:quadtree}.

\begin{figure}[ht]
    \includesvg[width=\textwidth]{images/invert-quadtree.svg}
    \caption{Visual representation of a quadtree.}
    \label{fig:quadtree}
\end{figure}

\subsection{The Stack and Heap}
A program uses two different types of memory, the stack and the heap. When a
program starts a fixed amount of bytes are allocated for it to use, this is what
is called the stack. When the program creates a variable it is allocated on the
stack, which is done by incrementing the stack pointer by the size of the
variable data type. Since a stack allocation is only a single instruction it is
incredibly fast. The heap is much larger than the stack but slower because you
have to request space within it from the operating system. Allocating is
considered to be a slow operation since there is a lot of overhead introduced.

\subsection{Big O notation}
Big O notation is a notation system to describe the time complexity of an
algorithm. If an algorithm has the time complexity of $O(n)$ then that means
its time complexity grows linearly with the amount of inputs $n$. This could be
deceiving due to a hidden constant within the expression. If $O(n)$ would be
written as a function to describe the actual execution time of the algorithm
then it would look like this $T(n) = C \cdot n$. A $O(n^{2})$ algorithm would
be considered worse than an $O(n)$ one. This assumption is, however, not always
correct. If $n$ is small enough then the $O(n^{2})$ algorithm could run faster
due to the constant being smaller than its $O(n)$ equivalent.

\subsection{Static Grid}
The static grid is built using a simple array. Indexing into the array is an
$O(1)$ operation. The caveat with this approach is you are forced to specify how
big the area you want the to partition to be before compiling the program. Both
insertion and querying only requires a simple index into the grid array which
makes this approach super fast.

\subsection{Hash Map}
A hash map is a data structure which stores key-value pairs allowing for
aliasing of values. For example, a hash map using strings as keys and integers
as values could map the key "life" to the value 42, meaning a query into the
hash map for "life" would return 42. The way this works is by using a
cryptographic hashing function to generate a unique hash for the object. This
hash is a simple number which is usually very large. To make the hash useful you
run modulo on it to make sure it's between zero and the number of buckets inside
the hash map. The same method is used for retrieval, first running the key
through the same hashing function, then wrapping it using modulo and returning
what lies in the bucket. The only caveat with this is there could be collisions
within the hash map. Two hashes could wrap into the same index. There are a
multitude of ways to handle collisions like open addressing and separate
chaining.

\subsection{Spatial Hashing Grid}
The spatial hashing grid is built using the same technique as a hash map. The
method I used to write my spatial hashing grid was to allocate an arbitrarily
sized array. When inserting into the structure I first hash the coordinate of
the object, then run a modulo operation on it giving me a number within the
array bounds, then insert it into the cell. No collision resolution is
taking place inside the structure meaning two cells could have the same bucket
index which means queries for one of the cells returns the objects in the other
as well.

\subsection{Vector}
A vector refers to a dynamic array, not the mathematical concept of a vector.
To allocate an array on the stack you need to know the size of it at compile
time which means vectors have to live on the heap. The vector works by having a
capacity which slowly fills up. Once it is full it resizes. This operation can
be quite expensive since it needs to request new memory from the operating
system and move all existing items into the new memory region.

\section{Material and Method}

\subsection{Linux}
Linux is the operating system used to develop the application and write this
paper. The program uses a specific Linux API to acquire the current time as a
UNIX time stamp. The way to do this would differ on other operating systems.

\subsection{Neovim}
Neovim was the editor used to edit the code. It is a modal text editor and not a
full integrated development environment, or IDE as they are called. A plugin
which utilizes the language server protocol was used to provide me with
intellisense and to aid in development by warning me about syntax errors.

\subsection{Git}
Git is a version control system which helps track changes to the project. In
this project it is utilized to keep a time line of the development process and
make sure changes could be synced between devices. The local git repository is
mirrored on GitHub making it accessible from other devices, and the source code
can be read from the browser.

\subsection{CMake}
CMake is a buildsystem, used to generate generate the final executable from the
source code. It aids in compiling and linking the program making cross
compilation and dependency management trivial. Since this is a simple project I
could have used a simple Makefile or even a shell script file but I opted for
CMake since it's industry standard. This, ultimately, had no effect on the
programs performance.

\subsection{C}
C was the program of choice for this project due to its low level control.
Since memory is managed manually in C I could be sure the results would not be
skewed by a garbage collector running in the middle of a benchmark. Its impact
is minimal and its fast build times allowed for quickly iterating during
development.

\subsection{SDL2}
SDL2 - Simple DirectMedia Layer - is a cross-platform library supplying an easy
API to create a window and render to it. This was crucial when developing since
visualizing the data structures made it easier to spot errors than simply
looking at the code. It also gives a nice visual representation of the program.
I used SDL2 instead of other similar libraries, such as Raylib, because I'm
familiar with it. Since the program is dealing with problems concerning two
dimensional space I found it easier to debug visually which is why SDL2 is a
dependency. This doesn't affect the results since it was only used during the
debugging phase of development.

\subsection{Python}
Python is an interpreted programming language which was used in a script to
generate graphs. All the graphs used in this paper was generated from data
collected from a C program, which dumped said data into a JSON file, which was
then read by a Python script using the matplotlib library to generate nice
looking graphs.

\section{Results \& Analysis}
This section presents the results and offers an analysis of the benchmarking
done on the three strategies - quadtrees, spatial hashing and static grids. The
metrics measured was average runtime: the sum of all operations needed to run
the program. Insertions: how long it took to insert all the objects into the
data structure. Collision testing: total time spent checking objects against
each other, which includes querying. Querying: average time spent retrieving
objects from the data structure. Clearing: how long it takes to reset the data
structure making it ready for rebuilding. Each metric was measured with a range
of object counts, labeled as 'objects' in the graphs, from 10 to 2500 with
increments of 10 objects between each measurement. At each stage was run 32 times
and the results were averaged to give smoother graphs. Two benchmarks was run,
one with an even distribution of objects and one with an uneven distribution.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{even-naive.png}
    \caption{Runtime of naive solution and partitioning strategies.}
    \label{fig:naive}
\end{figure}

Figure \ref{fig:naive} shows a drastic difference in runtime between the naive
solution and the optimization strategies. It is hard to evaluate which strategy
performs best due to the stark difference between the naive and the optimized
versions. Due to this discrepancy the naive solution will not be included in the
other graphs presented. This comparison was only run with 1000 objects due to
the time taken per execution.

\subsection{Average Runtime}

\subsubsection{Evenly Distributed Objects}
Figure \ref{fig:even-average} shows the average runtime for the program
using the different strategies with an even object distribution. This includes
all other measuring metrics. The X axis describes the amount of objects being
processed. The Y axis describes how much time was spent in milliseconds
executing the particular step in the program.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{even-average.png}
    \caption{Average runtime of spatial partitioning strategies with evenly
    distributed objects.}
    \label{fig:even-average}
\end{figure}

\subsubsection{Unevenly Distributed Objects}
Figure \ref{fig:uneven-average} presents the average runtime for the program
using the different strategies with an uneven object distribution. This includes
all other measuring metrics. As the object count increases the spatial hashing
trends towards a $O(n^2)$ runtime complexity.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{uneven-average.png}
    \caption{Average runtime of spatial partitioning strategies with unevenly
    distributed objects.}
    \label{fig:uneven-average}
\end{figure}

\subsubsection{Analysis}

With evenly distributed objects The three strategies perform very similarly,
though, spatial hashing does start showing more of an exponential runtime as the
object count increases. Could be due to the implementation of the spatial
hashing. Since it is using a hash map as its underlying data structure two
different cells could have the same index within the same hash map meaning
collision checks are happening for objects which are not close to each other in
space. This happens because there is a limited amount of buckets available to
store objects inside. The result of this scarcity is two hashes wrapping to the
same bucket index. As we can see in \ref{fig:uneven-average} this issue becomes
more apparent as the objects are less spread out.

\subsection{Insertion}

\subsubsection{Evenly Distributed Objects}
The runtime complexity of the insertion algorithms for evenly distributed
objects is shown in figure \ref{fig:even-insert}. The metric being times is the
insertion of all objects.
\begin{figure}[H]
    \centering
    \caption{Average insertion time per object with evenly distributed objects.}
    \includegraphics[width=0.5\textwidth]{even-insert.png}
    \label{fig:even-insert}
\end{figure}

\subsubsection{Unevenly Distributed Objects}
The runtime complexity of the insertion algorithms for evenly distributed
objects is shown in figure \ref{fig:uneven-insert}. The metric being times is
the insertion of all objects.
\begin{figure}[H]
    \centering
    \caption{Average insertion time per object with unevenly distributed
    objects.}
    \includegraphics[width=0.5\textwidth]{uneven-insert.png}
    \label{fig:uneven-insert}
\end{figure}

\subsubsection{Analysis}
There's virtually no difference in performance between the even and uneven
insertion times. This is because the codepath taken for each insertion is the
same, save for a different hash and bucket index. Because there is only a fixed
amount of objects able to be stored in each grid cell in both grid types the
insertion time is run in $O(1)$, constant time. The case isn't the same for the
quadtree since its a recursive algorithm. Each node has four sub-nodes which get
traversed each insertion leading to a $O(\log_4 n)$ runtime since each step in
the algorithm eliminates three of the sub-nodes. Once the correct node has been
found the insertion time is a constant $O(1)$ since it uses the same technique
as the other two.

\subsection{Querying}

\subsubsection{Evenly Distributed Objects}
Figure \ref{fig:even-query} represents the average time taken to query the space
per object with an even distribution.
\begin{figure}[H]
    \centering
    \caption{Average query time per object with evenly distributed objects.}
    \includegraphics[width=0.5\textwidth]{even-query.png}
    \label{fig:even-query}
\end{figure}

\subsubsection{Unevenly Distributed Objects}
Figure \ref{fig:uneven-query} represents the average time taken to query the space
per object with an uneven distribution.
\begin{figure}[H]
    \centering
    \caption{Average query time per object with unevenly distributed objects.}
    \includegraphics[width=0.5\textwidth]{uneven-query.png}
    \label{fig:uneven-query}
\end{figure}

\subsubsection{Analysis}
In both cases the quadtree query has almost exactly the same time with the same
$O(\log_4 n)$ runtime complexity. The other two, however, differ quite a bit
time wise. This discrepancy is due to the use of vectors. Since a vector is used
to return the overlapping objects the act of building this vector becomes more
expensive as the amount of objects increases. Since there are more objects near
each other in the unevenly distributed objects case there will naturally be more
objects added to the vector. Quadtrees do not suffer from the same weakness
since it has a limit on the amount of objects stored in each leaf meaning all
queries has an upper limit to the number of objects, eliminating the need for
vector resizing.

The reason for the jaggedness in the graphs is because the execution time is so
low the benchmark picks up on other things like the OS scheduler and other
application doing work in the background taking away resources from the
program.

\subsection{Collision}

\subsubsection{Evenly Distributed Objects}
Figure \ref{fig:even-collision} shows the total time spent checking for
colliding objects per iteration with an even distribution. This time also
includes the time taken for querying.
\begin{figure}[H]
    \centering
    \caption{Average collision time for all objects with evenly distributed
    objects.}
    \includegraphics[width=0.5\textwidth]{even-collision.png}
    \label{fig:even-collision}
\end{figure}

\subsubsection{Unevenly Distributed Objects}
Figure \ref{fig:uneven-collision} shows the total time spent checking for
colliding objects per iteration with an uneven distribution. This time also
includes the time taken for querying.
\begin{figure}[H]
    \centering
    \caption{Average collision time for all objects with unevenly distributed
    objects.}
    \includegraphics[width=0.5\textwidth]{uneven-collision.png}
    \label{fig:uneven-collision}
\end{figure}

\subsection{Analysis}
Just like with querying the same thing happens with collisions. Because there
are more objects close together there's a higher number of collision checks
taking place. Since the quadtree limit the amount of objects in a cell the
runtime doesn't change much between the two cases.

\section{Discussion}

\subsection{Design Choices}
The design choices for this project was entirely focused on CPU performance
leaving memory as an afterthought. This was done in order to mimic a modern game
running on consumer hardware where memory is often an abundant resource.
Everything inside the spatial partitioning strategies were allocated when the
structure was initialized so no heap allocations were being issued in the middle
of a hot loop. There are, however, two notable exception to this philosophy.
When running queries I used vectors, which have to allocate memory on the heap,
when doing queries. Looking back, there could be more efficient ways to handle
this. Instead of having the query function handle the memory management, let the
called handle that complexity. Have the caller allocate an array with sufficient
enough size and pass that as a parameter for the function to modify. Though this
would work it's not really a good solution, in my subjective opinion, because
there is guarantee of how many objects could be returned. The better solution
would be to use a bump, arena or block allocator to further speed up this
operation instead of using the general purpose malloc and free pattern provided
by the C standard library. The other notable exception is the benchmarking tool
I developed for this. It uses a hash map to store how many times a function was
run, the total time it ran and the average runtime per call. Because of my
underlying implementation of the hash map it will require memory allocation when
enough entries have been added. Not only that, but it was a nested structure
which created a new hash map for each benchmark scope.

\subsubsection{Spatial Hashing Grid}
No collision resolution was implemented for the spatial hashing grid because
there were very few collisions happening during testing so I didn't pay it much
mind. I thought the collision handling code would be more expensive than the
actual performance penalty from not handling the collisions. This was a naive
way of thinking. I should have implemented and measured instead of making such
bold assumptions.

\subsection{Compilation}
One thing I noticed after having generated all the graphs was that the program
was compiled in debug mode, meaning no compiler optimizations were enabled. If I
had changed the 'CMAKE\_BUILD\_TYPE' variable in my CMake build script from
'Debug' to 'Release' the results could be different due to the compiler doing
clever optimizations to speed up certain things. This could have lead to one
strategy out performing another. At the time of discovering this, it was too
late to explore this possibility further.

\subsection{Benchmarking}
The benchmarks are a thing I would have done differently if I would be redoing
this project. There are a few decisions that I can't justify now. One of these
weird choices is how I chose to benchmark insertions. Instead of running the
benchmark for each object inserted I measured to total time taken to insert all
objects. This doesn't sound weird until you look at how the rest of the
benchmarks are measured. The queries are measured once per object instead of
encapsulating all objects in a single benchmark. The insert metric is the only
one which differs in this regard. It would be an easy fix code wise, but having
to rewrite the paper based off of a new metric would be too much work to justify
the change.

\subsection{Pitfalls of Certain Strategies}
I found that although quadtrees outperform all the other strategies in this
particular program, it may not be the most suitable for all programs. Both grids
and quadtrees limit you to a fixed known world size which could be a problem for
larger games or simulations. If the quadtree is covering a massive region of the
world then that would introduce significant overhead because there would have to
be a lot of subdivisions happening. Chunking could fix this issue though there
are other strategies more suitable for this. The spatial hashing grid solves
this issue at the cost of performance. Grids come with the same world size restriction that the quadtree does but has
the benefit of being extremely trivial to implement.

\subsection{Use Cases}
As with almost all questions the answer to 'In what use case are the different
algorithms most useful?' the answer is: it depends. For static objects which
cover a known area, the quadtree is by far the most useful. For dynamic objects
which move a lot the spatial hashing grid is more suitable. And lastly, the
simple grid is really handy when you quickly want to increase the speed of a
system without having to spend a lot of time writing the optimization strategy,
which works really well for applications with a smaller amount of objects. A
combination of multiple strategies could be used to gain the benefit of them,
like utilizing the quadtree for static terrain while having all dynamic objects
be stored in a spatial hashing gird.

Then at what point is it worth implementing a more advanced data structure? Well
that also depends heavily on your particular use case. If the problem only
involves checking a handful of objects then the naive solution works just fine.
As seen in \ref{fig:naive} the naive solution is rarely viable as the object
count grows. The only real way to know when it is worth to implement a new
optimization strategy is to measure your current solution and make a decision
based off of that.

\section{Sources}
Nystom, Robert (2014) \textit{Game Programming Patterns}, genever benning

Nystom, Robert (2021) \textit{Crafting Interpreters}, genever benning

Stack Overflow (2012). What integer hash function are good that accepts an
integer hash key? Retrieved 2024-12-23, from
\url{https://stackoverflow.com/questions/664014/what-integer-hash-function-are-good-that-accepts-an-integer-hash-key}

\appendix

\section{Source Code}
The source code is shipped in a tarball with GZip compression name
'source.tar.gz'. Once unpacked all source code needed to compile and run it is
provided. If the SDL2 library isn't installed on the system then the code will
not compile.

To run the python script to generate the graphs python3 needs to be installed
alongside the matplotlib library.

A copy of the source code along side some other files can be found on the web
at \url{https://github.com/faintforge/spatial-partitioning}.

\end{document}